---
title: "hw01-daniel-vo.Rmd"
author: "Daniel Vo"
date: "2/21/2018"
output: github_document
---

# 1) Data Dictionary
# 2) Data Import
```{r}
library("ggplot2")
library("dplyr")
uci <- "https://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data"
download.file(
  url = uci, 
  destfile = "imports-85.data")

columnname <- c("symboling", "normalized_loss", "make", "fuel_type", "aspiration", "number_of_doors", 'body_style', 'drive_wheels', 'engine_location', 'wheel_base', 'length' , 'width', 'height' , "curb_weight", 'engine_type', 'num_of_cylinders', 'engine_system', 'fuel_system', 'bore', ' stroke' , 'compression_ratio', 'horse_power', 'peak_rpm', 'city_mpg', 'highway_mpg', 'price')
dt <- read.csv(file="~/imports-85.data", header = FALSE, col.names = columnname, stringsAsFactors = FALSE, na.strings = "")
dt$horse_power <- as.integer(as.character(dt$horse_power))
dt$peak_rpm <- as.integer(as.character(dt$peak_rpm))
dt$price <- as.numeric(as.character(dt$price))
str(dt)
```

# 3) Technical Questions about importing data 
a. If you don’t provide a vector of column names, what happens to the column names of
the imported data when you simply invoke read.csv('imports-85.data')?
ANSWER: By default, it would take the first row as the column names

b. If you don’t provide a vector of column names, what happens to the column names of the imported data when you invoke read.csv('imports-85.data', header = FALSE)?
ANSWER: There's no value. It would show "V1" , "V2", "V3", and so on.

c. When using the reading table functions, if you don’t specify how missing values are
codified, what happens to the data type of those columns that contain '?', e.g. price
or num_of_doors?
ANSWER: It would count "?"  in the data columns where data is character, and NA in the data columns where data is integer or numeric

d. Why is the data frame imported in the second option bigger (in terms of bytes) than the data frame imported in the first option?
ANSWER: I didn't actually remove the defaulted data type. I just put a new data type on top of the old one. So although it shows a different data type, it's just the name of the data type that is different

e. Say the object dat is the data frame produced when importing imports-85.data.
What happens to the data values if you convert dat as an R matrix?
ANSWER: Would become a vector with atomic values. I would not be able to create data frame with it. 

# 4) Practice base plotting
```{r}
library(ggplot2)
#histogram of price with colored bars
ggplot(data=dt, aes(dt$price)) + geom_histogram(aes(dt$price), fill=("red"))
# boxplot of horsepower in horizontal orientation.
boxplot(dt$horse_power, horizontal = TRUE, xlab="horse power")
# barplot of the frequencies of body_style, arranged in decreasing order. ***

ggplot(data=dt, aes(x = dt$body_style)) + geom_bar()

# stars() plot of vehicles with turbo aspiration, using only variables wheel-base, length, width, height, and price. ***

```

# 5) Summaries
```{r}
# What is the mean price of fuel_type gas cars? And what is the mean price of fuel_type diesel cars? (removing missing values) #gas = 12637.14 #diesel = 15838.15
d <- filter(dt, dt$fuel_type=="gas")
d$price[is.na(d$price)] <- 0
summarise(d, ave_price_gas = mean(price))
e <- filter(dt, dt$fuel_type=="diesel")
e$price[is.na(e$price)] <- 0
summarise(e, ave_price_diesel = mean(price))

# What is the make of the car with twelve num_of_cylinders? #jaguar
f <- filter(dt, dt$num_of_cylinder=="twelve")
factor(f$make)
# What is the make that has the most diesel cars? #mazda mercedes-benz nissan peugot toyota volkswagen volvo
factor(e$make)

# What is the price of the car with the largest amount of horsepower? #not available
dt$horse_power[is.na(dt$horse_power)] <- 0
dt$price[which.max(dt$horse_power)]

# What is the bottom 10th percentile of city_mpg? #12
summary(slice(filter(dt, dt$city_mpg<quantile(dt$city_mpg, probs = (0.1)))))

# What is the top 10th percentile of highway_mpg? #90
summary(slice(filter(dt, dt$highway_mpg>quantile(dt$city_mpg, probs = (0.9)))))

# What is the median price of those cars in the bottom 10th percentile of city_mpg? #6575
h <- slice(filter(dt, dt$city_mpg>quantile(dt$city_mpg, probs = (0.9))))
h$price[is.na(h$price)] <- 0
summarise(h, median = median(h$price))
```

# 6) Technical Questions about data frames
a. What happens when you use the dollar $ operator on a data frame, attempting to use
the name of a column that does not exist?
ANSWER: showing NULL

b. Which of the following commands fails to return the vector mpg which is a column in
the built-in data frame mtcars:
ANSWER: (4) mtcars[ ,mpg]

c. Based on your answer for part (b), what is the reason that makes such command to
fail?
ANSWER: mpg is not an object within the data frame. It is the name of a vector in a column

d. Can you include an R list as a “column” of a data frame? YES or NO, and why.
ANSWER: YES. But I have to convert it into a data frame, then allow it to merge in with the existing data frame

e. What happens when you apply as.list() to a data frame? e.g. as.list(mtcars)
ANSWER: It would show all the vectors associated with each of the variables in the data frame. The order of the elements in each vector follow the rows order in the data frame. 

f. Consider the command: abc <- as.list(mtcars). What function(s) can you use to
convert the object abc into a data frame?
ANSWER: Function unlist(abc) to make vector
Function matrix() to create a matrix with nrow (#of rows) and ncol (#of columns)
Function data.frame to create a data frame

# 7) Correlations of quantitative variables
```{r}
install.packages("corrplot")
library("corrplot")

qdat <- subset(dt, select = -c(symboling, normalized_loss, make, fuel_type, aspiration, number_of_doors, body_style, drive_wheels, engine_location, engine_type, num_of_cylinders, fuel_system, bore, X.stroke) )
qdat <- na.omit(qdat)
qdat
correl <- cor(qdat)
corrplot(correl, method="pie")
corrplot(correl, method="color")
##Comment:
#I see a lot of correlation between variables. While roughly 2/3 of the correlation is positive, and 1/3 is negative. The degree of correlation varies 
```

# 8) Principal Components Analysis

```{r}
qdat1 <- princomp(qdat, cor = TRUE)
eigenvalues <- qdat1$sdev^2
eigenvalues
pror_qdat <- eigenvalues/sum(eigenvalues)
data.frame(pror_qdat)

qdat1$loadings
biplot(qdat1, scale = 0)

#Comment: Based on the graph, we see peak_rpm and compression_ratio are the two components that have the most variability. 

```











